---
title: "Instructor Pay 2026"
format: html
knitr:
  opts_chunk: 
    collapse: true
    comment: "#" 
    echo: false
    message: false
---

```{r setup}
library(googlesheets4)
library(ggplot2)
library(dplyr)
library(lubridate)
library(viridis)
library(forcats)
library(plotly)
library(DT)
library(purrr)
library(rpc)
```

```{r}
class_sheet <- "https://docs.google.com/spreadsheets/d/1vivUrj8WSWI2hHTOlgdWfiHdyTRd0zCEIX9xJnHAf54/edit?gid=1058139562#gid=1058139562"
sessions <- c("2025 Fall", "2026 Winter", "2026 Spring")

class_schedule <- map(sessions, ~ read_sheet(class_sheet, sheet = .x))

names(class_schedule) <- sessions
```

```{r}
calculate_class_revenue <- function(df, type, enrolled) {
  class_df <- filter(df, class_type == type) |>
    mutate(total_revenue = cost * enrolled)

  sum(class_df$total_revenue)
}
```

## Previous work
A preliminary break down of class costs can be found in the [google sheet](https://docs.google.com/spreadsheets/d/1-EH4VNp-e8omI9PokcXmci-RPvE_Omm7u9u16iKgQ5Q/edit?pli=1&gid=0#gid=0)

The [instructor pay analysis in comparable studios](https://docs.google.com/document/d/1EjXD1VmmGHBGLcv-ICNnXFTIaojOH6x5JEzzPV9auXA/edit?tab=t.xstvt0mq2lke#heading=h.zbl0ep6q3q0u) done by Crystal Yee - with ChatGPT assistance

- I'm unsure how comparable these rates are or how they compare to our current instructors. 

I believe there was also a class comparison some where as well but I do not have access to it.

## Public classes
```{r}
#map(class_schedule, ~ calculate_class_revenue(.x, "public", 10))
other_expenses <- 1874
```

```{r calculations}
instructors_base <- data.frame(
  name = c("Jean Fung", "Evan Leung", "David Liu"),
  number_of_courses = c(1, 3, 3),
  rate = c(30, 30, 30)
)

instructors <- instructors_base

for (raise in 1:20) {
  instructors <- rbind(
    instructors,
    instructors_base |> mutate(rate = rate + raise)
  )
}

instructor_pay <- instructors |>
  rowwise() |>
  mutate(
    pay = calculate_instructor_pay(
      rate,
      "public",
      number_of_courses,
      name
    )
  )

revenue <- data.frame(
  class_price = seq(from = 250, to = 350, by = 25)
) |>
  mutate(
    total = 70 * class_price,
    after_city_cut = total * 0.75,
    after_other_expenses = after_city_cut - other_expenses
  )

projected_pay <- instructor_pay |>
  group_by(rate) |>
  summarise(total_instructor_compensation = sum(pay))

percent_revenue <- projected_pay
for (class_price in revenue$class_price) {
  class_income <- (class_price * 70)

  percent_revenue <- percent_revenue |>
    mutate(
      "profit_{{class_price}}" := (class_income * 0.75) -
        other_expenses -
        total_instructor_compensation,
      "revenue_percent_{{class_price}}" := round(
        ((class_income -
          total_instructor_compensation) /
          class_income *
          100),
        1
      )
    )
}

# percent_revenue_after_city <- projected_pay
# for (class_price in revenue$class_price) {
#   class_increase <- (class_price * 70) * 0.75

#   percent_revenue_after_city <- percent_revenue_after_city |>
#     mutate(
#       "revenue_percent_{{class_price}}" := ((class_increase -
#         total_instructor_compensation) /
#         class_increase) *
#         100
#     )
#}
```

The projected pay for each dollar increase in pay:

- The current instructor pay rate is at $36/hour.
- The previous pay increase was $30 (4-5 years ago)
- Keep in mind during this time there was considerable cost of living increases and inflation. A estimate of % increase per year can be found - Cost of living example - https://mpp.pensionsbc.ca/cost-of-living-adjustments
- Instructors are not paid vacation or any other benefits and will need to pay any other deductions on their own.
- We do provide clay for their use during class.
```{r}
projected_pay |>
  datatable()
```


Public classes are 9 classes for $275 per students in 2026 an increase of  $25 from $250 from previous years. We are currently running 7 sessions of 10 students each totaling 70 students per session.

For an increase in the class cost per student (`class_price`). The `total` expected income and deductions `after_city_cut` of 25% and `after_other expenses` (glaze, square fees, insurance, clay etc.)
```{r}
revenue |>
  datatable()
```

Table of the instructor expenses versus the percentage of revenue (before any deductions). I would use this to determine the approximate ratio that we want to keep the instructor pay at and adjust class prices accordingly so we make a reasonable profit with future increases.

I've included previous prices ($250) and rates ($30) as a point of comparison.

I would suggest at least a $4 increase to $40 and also consider a increase in class costs to off set the increased cost.
```{r}
percent_revenue |>
  datatable()
```


## Member classes
Pay increase for member classes by a 1 dollar per hour increase:
```{r}
#map(class_schedule, ~ calculate_class_revenue(.x, "member", 10))

instructors_base <- data.frame(
  name = c("David Liu"),
  number_of_courses = c(1),
  rate = c(36)
)

instructors <- instructors_base

for (raise in 1:10) {
  instructors <- rbind(
    instructors,
    instructors_base |> mutate(rate = rate + raise)
  )
}

instructors |>
  rowwise() |>
  mutate(
    pay = calculate_instructor_pay(
      rate,
      "member",
      number_of_courses,
      name
    )
  ) |>
  select(-name) |>
  datatable()
```

Member classes have operated at more or less a just breakeven model to make sure we cover our costs. It is more of a member benefit.
```{r}
data.frame(
  class_price = seq(from = 165, to = 250, by = 10)
) |>
  mutate(
    total = 10 * class_price,
    after_other_expenses = total - 1102.4
  ) |>
  datatable()
```

### Follow up

- Because we have kept prices the stagnant for several years we have some adjustment to do and we may need to introduce it gradually.
-  Annual class cost and pay review to make sure we keep up with current costs
- Make sure we keep up with Instructors