---
title: "Member Demographics"
subtitle: "Welcome Back"
execute:
  echo: false
  warning: false
  message: false
---

```{r library}
library(googlesheets4)
library(dplyr)
library(ggplot2)
library(scales)
library(janitor)
library(plotly)
library(forcats)
```

```{r data}
sheet <- "https://docs.google.com/spreadsheets/d/1Jt1LKbq5DxXn8p7QOAUzR0_B2FfIXRNhwevpwNf6nXE/edit?gid=0#gid=0"

raw_renew <- read_sheet(ss = sheet, sheet = "Raw Members Renewing") |>
  clean_names()

raw_new <- read_sheet(ss = sheet, sheet = "Raw New Members") |>
  clean_names()
```

## Member Space Access
```{r frequency}
studio_frequency <- raw_renew |>
  count(in_studio_frequency)

frequency_levels <- c(
  "Less than once per week",
  "1-2 times per week",
  "3 or more times per week"
)

studio_plot <- studio_frequency |>
  ggplot(aes(x = in_studio_frequency, y = n)) +
  geom_col(fill = "#2C6E91", width = 0.7) +
  scale_x_discrete(limits = frequency_levels) +
  labs(
    title = "How often are members using the studio",
    subtitle = paste0(
      "Share of ",
      nrow(raw_renew),
      " renewing members"
    ),
    x = NULL,
    y = "number of members"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
  )

ggplotly(studio_plot)
```

```{r usage}
studio_new <- raw_new |>
  count(home_studio) |>
  mutate(member_status = "new")

studio_renew <- raw_renew |>
  count(home_studio) |>
  mutate(member_status = "renew")

usage_plot <- bind_rows(studio_new, studio_renew) |>
  mutate(home_studio = fct_reorder(home_studio, n, .fun = sum)) |> # order bars by total
  ggplot(aes(x = home_studio, y = n, fill = member_status)) +
  geom_col(position = "stack") +
  coord_flip() + # horizontal bars: labels easy to read
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_fill_manual(
    values = c("new" = "#2C7BB6", "renew" = "#F28E2B"),
    labels = c("new" = "New", "renew" = "Renewing Memebers"),
    name = "Member status"
  ) +
  labs(
    title = "Members by Home Studio",
    x = "",
    y = "Number of members"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text = element_text(color = "black"),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold", size = 16),
    legend.position = "top",
    panel.grid.major.y = element_blank()
  )

ggplotly(usage_plot)
```

```{r}
tenure_summary <- raw_renew |>
  count(membership_history) |>
  arrange(n)
```

## Membership Composition

General Takeaways
Caring and service professions dominate. Education and healthcare together make up roughly a third of the group.
A strong creative and design contingent sits alongside a solid technical and business contingent.
Mixed life stages. There are several retirees, people on maternity leave, full-time parents or homemakers, and current students, as well as people mid-career.
Public sector connections (government, City of Richmond, WorkSafeBC, UBC, KPU, schools) are common.
```{r}
# install.packages("networkD3")
library(networkD3)

nodes <- data.frame(
  name = c(
    "Prior members (91)",
    "Renewed (84)",
    "Not renewing (7)",
    "New (18)",
    "Current membership (102)"
  )
)
links <- data.frame(
  source = c(0, 0, 1, 3),
  target = c(1, 2, 4, 4),
  value = c(84, 7, 84, 18)
)

my_colors <- 'd3.scaleOrdinal()
  .domain(["Total membership (109)","New (18)","Renewing (84)","Not renewing (7)"])
  .range(["#999999","#2E86AB","#3BB273","#E15554"])'

sankeyNetwork(
  Links = links,
  Nodes = nodes,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "name",
  colourScale = my_colors,
  fontSize = 14,
  nodeWidth = 30
)
```

## Length of Membership

```{r}
#| fig-width: 9
#| fig-height: 5

tenure_levels <- c(
  "Less than 5 years",
  "Between 5-10 years",
  "More than 10 years",
  NA
)

tenure_summary |>
  ggplot(aes(x = membership_history, y = n)) +
  geom_col(fill = "#2C6E91", width = 0.7) +
  scale_x_discrete(limits = tenure_levels) +
  labs(
    title = "How long have our members been with the club?",
    subtitle = paste0(
      "Share of ",
      nrow(raw_renew),
      " members by length of membership"
    ),
    x = NULL,
    y = "number of members"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
  )
```