---
title: "Studio Use"
format: html
knitr:
  opts_chunk: 
    collapse: true
    comment: "#" 
    echo: false
    message: false
---

```{r setup}
library(googlesheets4)
library(ggplot2)
library(dplyr)
library(lubridate)
library(viridis)
library(forcats)
library(plotly)
library(DT)
library(glue)
library(tidyr)
```

```{r google_sheet}
membership_sheet <- "https://docs.google.com/spreadsheets/d/1Ldrcx_iVWi2Bv3GQoIHE2hOkf0Ah-nKVK3GFnXBXfts/edit?gid=1039957161#gid=1039957161"
membership <- read_sheet(membership_sheet, sheet = "2025-2026")

sheet <- "https://docs.google.com/spreadsheets/d/1-XhBYaeJI-uwPJsDb2ucxgnifXC1WdGjejI9a0QK0Po/edit?gid=1184437612#gid=1184437612"
signin <- read_sheet(sheet)
```

```{r clean_up}
signin_clean <- signin |>
  select(Timestamp, `Column 2`, Action) |>
  mutate(
    date = date(Timestamp),
    month = month(date),
    day_of_week = wday(date, label = TRUE)
  ) |>
  rename("member" = "Column 2")
```

```{r signout_analysis}
no_signout <- signin_clean |>
  group_by(member, date) |>
  count() |>
  filter(n == 1) |>
  nrow()

multi_signin <- signin_clean |>
  group_by(member, date) |>
  count() |>
  mutate(in_studio = n / 2) |>
  filter(in_studio > 1)
```

### Data Limitations

-   During this time members still signed in manually
-   Some new member sign ins were not captured
-   Not all sign out times were logged
-   `r no_signout` of `r NROW(signin_clean)` (`r round((no_signout / NROW(signin_clean)) * 100, 2)` %) entries did not have a sign out

### Findings

```{r usage}
people_per_day <- signin_clean |>
  filter(Action == "Sign In") |>
  group_by(date) |>
  count() |>
  mutate(month = month(date), day_of_week = wday(date, label = TRUE))
```

```{r}
member_usage <- signin_clean |>
  group_by(member) |>
  count() |>
  arrange(desc(n))
```

### By Person usage

-   Only captures people using the sign in form
-   `r NROW(member_usage)`/ `r NROW(membership)` (`r round((NROW(member_usage)/NROW(membership))*100, 2)` %) has been using the sign in form or coming into the studio

#### Studio usage over days of the week

```{r calendar_plot}
day_levels <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")

closed <- data.frame(
  date = as.Date(c("2025-12-25", "2025-12-26", "2026-01-01")),
  n = NA_integer_
) |>
  mutate(
    week_start = floor_date(date, unit = "week", week_start = 1),
    day_of_week = wday(date, label = TRUE)
  )

people_per_date_formatted <- people_per_day |>
  filter(date > "2025-11-27") |>
  mutate(week_start = floor_date(date, unit = "week", week_start = 1)) |>
  select(date, week_start, day_of_week, n)

plot_data <- bind_rows(people_per_date_formatted, closed) |>
  mutate(
    doy = factor(day_of_week, levels = rev(day_levels)),
    tooltip_text = glue(
      "{format(date, '%a, %b %d %Y')}\nSign Ins: {ifelse(is.na(n), 'Closed', n)}"
    )
  )

plot <- plot_data |>
  ggplot(aes(x = week_start, y = doy, fill = n, text = tooltip_text)) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_viridis(name = "Sign Ins", na.value = "grey85") +
  scale_x_date(
    date_breaks = "1 week",
    date_labels = "%b %d",
    expand = c(0, 0)
  ) +
  labs(title = "Studio Sign Ins", x = "Week of", y = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 14)
  )

ggplotly(plot, tooltip = "text") |>
  layout(hoverlabel = list(align = "left"))
```

```{r member_hours}
duplicate_entry <- signin_clean |>
  dplyr::summarise(n = dplyr::n(), .by = c(member, date, Action)) |>
  dplyr::filter(n > 1L)

member_time <- signin_clean |>
  anti_join(duplicate_entry, by = join_by(member, date, Action)) |>
  pivot_wider(names_from = Action, values_from = Timestamp) |>
  filter(!is.na(`Sign In`)) |>
  filter(!is.na(`Sign Out`)) |>
  mutate(
    month_start = floor_date(date, unit = "month"),
    duration = time_length(interval(`Sign In`, `Sign Out`), unit = "minute")
  ) |>
  filter(duration > 10) |>
  group_by(month_start) |>
  summarise(avg_time_spent = round(mean(duration) / 60, 1))
```

#### Average Studio Use Length
-   Members that did not sign in or out is excluded
-   Assuming the Time stamp time was sign in sign out time
-   Filtered out entries where there are multiple entries
-   Average member use time `r mean(member_time$avg_time_spent)` hours
-   This is constrained by classes in session and studio availability

#### Average Member use per month

```{r monthly_summary}
monthly_usage <- people_per_day |>
  mutate(month_start = floor_date(date, unit = "month")) |>
  group_by(month_start) |>
  summarise(
    total_sign_ins = sum(n, na.rm = TRUE),
    days_with_activity = n_distinct(date),
    avg_sign_ins_per_day = round(total_sign_ins / days_with_activity, 1),
    .groups = "drop"
  ) |>
  arrange(month_start) |>
  mutate(
    Month = format(month_start, "%B %Y"),
  ) |>
  left_join(member_time, by = join_by(month_start == month_start)) |>
  select(
    Month,
    `Total Sign Ins` = total_sign_ins,
    `Days Active` = days_with_activity,
    `Average number of sign ins per Active Day` = avg_sign_ins_per_day,
    `Average number of hours spent` = avg_time_spent
  )

datatable(
  monthly_usage,
  rownames = FALSE,
  options = list(
    dom = "t",
    ordering = FALSE,
    pageLength = nrow(monthly_usage)
  ),
  caption = "Studio Usage by Month"
) |>
  formatStyle(
    "Total Sign Ins",
    background = styleColorBar(
      range(monthly_usage$`Total Sign Ins`),
      "lightblue"
    ),
    backgroundSize = "98% 88%",
    backgroundRepeat = "no-repeat",
    backgroundPosition = "center"
  ) |>
  formatStyle(
    "Average number of sign ins per Active Day",
    background = styleColorBar(
      range(monthly_usage$`Average number of sign ins per Active Day`),
      "#fdd49e"
    ),
    backgroundSize = "98% 88%",
    backgroundRepeat = "no-repeat",
    backgroundPosition = "center"
  ) |>
  formatStyle(
    "Month",
    fontWeight = "bold"
  )
```
