---
title: "Richmond Potters' Club — Volunteer Hours Report"
subtitle: "Annual General Meeting `r format(Sys.Date(), '%B %d, %Y')`"
format:
  html:
    toc: true
    toc-title: "Report Sections"
    toc-depth: 2
    theme: cosmo
    css: styles.css
    embed-resources: true
    page-layout: full
knitr:
  opts_chunk:
    collapse: true
    comment: "#"
    echo: false
    message: false
    warning: false
---

```{r setup}
library(googlesheets4)
library(ggplot2)
library(dplyr)
library(lubridate)
library(DT)
library(glue)
library(tidyr)
library(plotly)
library(stringr)
library(purrr)
```

```{r google_sheet}
membership_sheet <- "https://docs.google.com/spreadsheets/d/1Ldrcx_iVWi2Bv3GQoIHE2hOkf0Ah-nKVK3GFnXBXfts/edit?gid=1039957161#gid=1039957161"
membership <- read_sheet(membership_sheet, sheet = "2025-2026")

volunteer_sheet_url <- "https://docs.google.com/spreadsheets/d/1Tj7O8SzHRy_FgNHE84TD9od4CIbv_WVIpoqORUdNB8s/edit?gid=303571686#gid=303571686"
volunteer_sheet <- read_sheet(volunteer_sheet_url, sheet = "Total Raw Data")
```

```{r data_prep}
ordered_year_month <- map(
  2025:year(Sys.Date()),
  ~ paste(.x, month.abb)
) |>
  unlist()

volunteer_sheet_clean <- volunteer_sheet |>
  mutate(
    month = month(`Date of Volunteer Activity`, label = TRUE),
    year = year(`Date of Volunteer Activity`),
    year_month = glue("{year} {month}"),
    year_month_factor = factor(year_month, levels = ordered_year_month),
    member = str_replace_all(`Column 2`, "\\b([A-Za-z])[A-Za-z]*\\b", "\\1")
  ) |>
  filter(`Column 2` != "Justin Ng") |>
  select(-Branch) |>
  left_join(membership, by = join_by("Column 2" == "Full Name"))

branch_members <- membership |>
  count(`Assigned Branch`)

exempt <- branch_members |>
  filter(`Assigned Branch` == "Exempt") |>
  pull(n)

# Summary stats used across sections
total_hours <- sum(volunteer_sheet_clean$Hours, na.rm = TRUE)
total_members <- nrow(membership)
hours_threshold <- 34 # minimum required hours

volunteer_totals <- volunteer_sheet_clean |>
  group_by(`Column 2`) |>
  summarise(total_hours = sum(Hours, na.rm = TRUE))

members_over <- sum(volunteer_totals$total_hours > hours_threshold)
members_under <- sum(volunteer_totals$total_hours <= hours_threshold)
pct_compliant <- round(100 * members_over / nrow(volunteer_totals))
```

---

## 📊 Key Highlights {#highlights}

> *Data as of `r format(Sys.Date(), "%B %d, %Y")`*

These three numbers tell the top-level story for the `r format(Sys.Date(), "%Y")` membership year where there are `r exempt` of `r total_members`.

```{r kpi_cards}
# Inline KPI summary rendered as an HTML table for visual emphasis
kpi_html <- glue(
  '
<div style="display: flex; gap: 1.5rem; flex-wrap: wrap; margin: 1rem 0 2rem 0;">

  <div style="flex:1; min-width:180px; background:#2c5f2e; color:white;
              border-radius:12px; padding:1.5rem; text-align:center;">
    <div style="font-size:2.8rem; font-weight:700; line-height:1;">
      {format(round(total_hours), big.mark=",")}
    </div>
    <div style="font-size:1rem; margin-top:0.4rem; opacity:0.85;">
      Total Hours Contributed
    </div>
  </div>

  <div style="flex:1; min-width:180px; background:#1a3a5c; color:white;
              border-radius:12px; padding:1.5rem; text-align:center;">
    <div style="font-size:2.8rem; font-weight:700; line-height:1;">
      {pct_compliant}%
    </div>
    <div style="font-size:1rem; margin-top:0.4rem; opacity:0.85;">
      Members Meeting {hours_threshold}-Hour Minimum
    </div>
  </div>

  <div style="flex:1; min-width:180px; background:#7a4f1e; color:white;
              border-radius:12px; padding:1.5rem; text-align:center;">
    <div style="font-size:2.8rem; font-weight:700; line-height:1;">
      {round(total_hours / total_members, 1)}
    </div>
    <div style="font-size:1rem; margin-top:0.4rem; opacity:0.85;">
      Average Hours Per Member
    </div>
  </div>

</div>
'
)

htmltools::HTML(kpi_html)
```

::: {.callout-tip title="What this means for members"}
A strong year overall — thank you to everyone who contributed! See below for the branch breakdown and to check your individual standing against the `r hours_threshold`-hour minimum.
:::

---

## 🏛️ Hours by Branch {#by-branch}

### Branch Summary Table

Use this table to see how your branch performed relative to its size.

```{r branch_table}
branch_summary <- volunteer_sheet_clean |>
  filter(!(`Assigned Branch` %in% c("Exempt")), !is.na(`Assigned Branch`)) |>
  group_by(`Assigned Branch`) |>
  summarise(`Total Branch Hours` = round(sum(Hours, na.rm = TRUE))) |>
  left_join(branch_members, by = "Assigned Branch") |>
  mutate(
    `Average Hours Per Member` = round(`Total Branch Hours` / n, 1),
    `Members` = n
  ) |>
  select(
    `Assigned Branch`,
    Members,
    `Total Branch Hours`,
    `Average Hours Per Member`
  ) |>
  arrange(desc(`Assigned Branch`))

datatable(
  branch_summary,
  options = list(pageLength = 10, dom = "t"),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: bottom; text-align: left; color: #666;",
    "Hours per member uses current membership count as denominator."
  )
)
```

### Monthly Hours by Branch

The chart below shows contribution trends over time. Hover over bars for details.

```{r branch_monthly_chart}
branch_monthly <- volunteer_sheet_clean |>
  filter(!(`Assigned Branch` %in% c("Exempt")), !is.na(`Assigned Branch`)) |>
  group_by(year_month_factor, `Assigned Branch`) |>
  summarise(total_hours = sum(Hours, na.rm = TRUE), .groups = "drop")

p_stacked <- ggplot(
  branch_monthly,
  aes(
    x = year_month_factor,
    y = total_hours,
    fill = `Assigned Branch`,
    text = paste0(
      `Assigned Branch`,
      "<br>",
      year_month_factor,
      ": ",
      round(total_hours),
      " hrs"
    )
  )
) +
  geom_col() +
  labs(
    x = NULL,
    y = "Hours",
    fill = "Branch",
    title = "Monthly Volunteer Hours — All Branches"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold"),
    legend.position = "bottom"
  )

ggplotly(p_stacked, tooltip = "text") |>
  layout(legend = list(orientation = "h", y = -0.25))
```

### Per-Branch Trend (Faceted)

```{r branch_facet_chart}
p_facet <- ggplot(
  branch_monthly,
  aes(
    x = year_month_factor,
    y = total_hours,
    fill = `Assigned Branch`,
    text = paste0(
      year_month_factor,
      ": ",
      round(total_hours),
      " hrs"
    )
  )
) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~`Assigned Branch`, scales = "free_y") +
  labs(x = NULL, y = "Hours", title = "Monthly Hours per Branch") +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold")
  )

ggplotly(p_facet, tooltip = "text")
```

---

## 👤 Individual Member Hours {#individual}

### Are You Meeting the Minimum?

The `r hours_threshold`-hour minimum is required for membership in good standing.
Currently **`r members_over` of `r nrow(volunteer_totals)` members** (`r pct_compliant`%) have met or exceeded this threshold.

```{r individual_chart}
volunteer_totals_plot <- volunteer_sheet_clean |>
  group_by(`Column 2`) |>
  summarise(total_hours = sum(Hours, na.rm = TRUE)) |>
  arrange(total_hours) |>
  mutate(
    member_factor = factor(`Column 2`, levels = `Column 2`),
    status = ifelse(
      total_hours >= hours_threshold,
      "Met minimum",
      "Below minimum"
    ),
    label_text = paste0(`Column 2`, ": ", round(total_hours), " hrs")
  )

p_individual <- ggplot(
  volunteer_totals_plot,
  aes(x = member_factor, y = total_hours, fill = status, text = label_text)
) +
  geom_col() +
  geom_hline(
    yintercept = hours_threshold,
    linetype = "dashed",
    colour = "black",
    linewidth = 0.6
  ) +
  annotate(
    "text",
    x = 1,
    y = hours_threshold + 2,
    label = glue("{hours_threshold}-hr minimum"),
    hjust = 0,
    size = 3.2,
    colour = "black"
  ) +
  scale_fill_manual(
    values = c(
      "Met minimum" = "#2c7a2e",
      "Below minimum" = "#c0392b"
    )
  ) +
  labs(
    x = NULL,
    y = "Total Hours",
    fill = NULL,
    title = "Cumulative Volunteer Hours per Member"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.title = element_text(face = "bold"),
    legend.position = "bottom"
  )

p_individual
```

::: {.callout-warning title="Action required — members below minimum"}
**`r members_under` members** have not yet reached `r hours_threshold` hours.
If you are below the line, please contact your branch coordinator to find upcoming volunteer opportunities before the membership year closes.
:::

*Report generated automatically from club records on `r format(Sys.Date(), "%B %d, %Y")`.
Contact the club secretary with any data discrepancies.*
